### 2017-03-20 intro.tex
intro finalized

%!TEX root =  autocontgrlp.tex

>>>>

%!TEX root =  autocontgrlp.tex
\section{Introduction}
Markov decision processes (MDPs) have proved to be an indispensable model for sequential decision making under uncertainty with applications in networking, traffic control, robotics, operations research, business, finance, artificial intelligence, health-care and more (see, e.g., \cite{
White93:Apps,
rust96:book,
FeiSh02:MDPHandbook,
QiWu07,
SiBu10:MDPinAI,
BauRie:11,Puter,
LeLiu12:RLBook,
Abuetal15:MDPWireless,
BouDi17:MDPPractice}).
In this paper we adopt the framework of discrete-time, discounted MDPs when
a controller steers the stochastically evolving state of a system while receiving 
rewards that depends on the states visited and actions chosen. The goal is to choose the actions so as to maximize the \emph{return}, defined as the total discounted expected reward. A controller that uses past state information is called a \emph{policy}. An \emph{optimal policy} is one that maximizes the value no matter where the process is started from \cite{Puter}.
In this paper we consider planning problems where the goal is to calculate actions of policies that give rise to high values
 and give new error bounds on the quality of solutions obtained by solving linear programs of tractable size. 
To explain the contributions in more details, we start by describing the computational challenges involved in planning.

The main objective of \emph{planning} is to compute actions of an optimal policy while interacting with an MDP model. 
In finite state-action MDPs,
assuming access to individual transition probabilities and rewards along transitions, 
various algorithms are available to perform this computation in time and space that scales polynomially with the number of states and actions.
However, in most practical applications, the MDP is compactly represented
and if it is not infinite, the number of states scale \emph{exponentially} with the \emph{size of the representation} of the MDP.
%an effect that is known as \emph{Bellman's curse of dimensionality}.
If planners are allowed to perform some fixed amount of calculations for each state encountered,
it is possible to use sampling to make the per-state calculation-cost 
independent of the size of the state space \cite{rust96:randomization,szepesvari2001,kearns2002sparse}.
Nevertheless, the resulting methods are still quite limited. 
In fact, various hardness results show that computing actions of (near-) optimal policies is intractable in various senses
and in various compactly represented MDPs \cite{BlonTsi:00Complexity}.
Given these negative results, 
it is customary to adopt the \emph{modest goal of efficiently computing actions of a policy that 
is nearly as good as the best policy in a given restricted policy class}.
Here, within some restrictions (see below), the policy class can be chosen by the user.
The more flexibility the user is given in this choice, the stronger a planning method is.

A popular approach along these lines, which goes back to \citet{SchSei85}, 
relies on considering linear approximations to the \emph{optimal value function}:
The idea is that, similarly to linear regression, a fixed sequence of basis functions are combined
linearly. The user's task is to use a priori knowledge of the MDP 
to choose the basis functions so that 
a good approximation to the optimal value function will exist in the linear space spanned by the basis functions.
The idea then is to design some algorithm to find the coefficients of the basis functions that gives a a good approximation, 
while keeping computation cost in check.
Finding a good approximation is sufficient, since at the expense of an extra $O(1/\epsilon^2)$ randomized computation, 
a uniform $O(\epsilon)$-approximation 
to the optimal value function can be used to calculate an action of an $O(\epsilon)$-optimal policy at any given state
(e.g., follow the ideas in \cite{szepesvari2001,kearns2002sparse}; see also Theorem 3.7 of \citet{Kall17}).
Since the number of coefficients can be much smaller than the number of states, the algorithms that search
for the coefficients have the potential to run efficiently regardless of the number of states.

Following \citet{SchSei85}, most of the literature considers
algorithms that are obtained from restricting exact planning methods to search 
in the span of the fixed basis functions when performing computations.
In this paper we consider the so-called \emph{approximate linear programming} (ALP) approach, which
was heavily studied during the last two decades 
\cite[e.g.,][]{
schuurmans,
gkp,
ALP,
CS,
kveton2004heuristic,
petrik,
SALP,
fs,
npalp,
BhatFaMo12:SALPNP,
abbasi}.
The basic idea here is to combine a linear program whose solution is the optimal value function (and thus the number of optimization variables in it scales with the number of states) with a linear constraint that restricts the optimization variables to lie in the subspace spanned by the basis functions. As already noted by \citet{SchSei85}, the new LP can still be kept feasible by just adding one special basis function, while by substituting the ``value function candidates'' with their linear expansions, the number of optimization variables becomes the number of basis functions. 
As shown by \citet{ALP}, the solution to the resulting LP is within a constant factor of the best approximation to the optimal value function within the span of the chosen bases. However, since the number of constraints in the LP is still proportional to the number of states, it is not obvious whether a solution to the resulting LP can be found in time independent of the number of states (other computations can be done in time independent of the number of states, e.g., using sampling, at the price of a controlled increase of the error, e.g., Theorem 6 of \citep{petrik}).

Most of the literature is thus devoted to designing methods to select a tractable subset of the constraints while keeping the approximation guarantees, as well as keeping computations tractable. 
For example, \citet{schuurmans} and \citet{gkp} propose constraint generation in a setting 
where the MDP has additional structure (i.e., factorized transition structure) 
that is then exploited in the constraint generation method.
A more general approach due to \citet{CS} 
is to choose a random subset of the constraints 
by choosing states to be included at random from a distribution that reflects the ``importance'' of states.
While constraint generation is powerful when applicable,
it is not known how solution quality degrades with the budget on the constraints generated.
For constraint sampling,  \citet{CS} prove a bound on the suboptimality, but this bound applies only 
in the unrealistic scenario when the constraints are sampled from an \emph{idealized} distribution,
which is related to the stationary distribution of an optimal policy. 
While it is possible to extend this result
to any sampling distribution, the bound then scales with the mismatch between the sampling and the idealized
distributions, which, in general, will be uncontrolled. 
Another weakness of the bound is related to that when constraints are dropped, the linear program may become
unbounded. To prevent this, \citet{CS}  propose adding an extra constraint on the optimization variables.
The bound they obtain, however, scales with the \emph{worst approximation error} over this constraint set.
While in a specific example it is shown that this error can be controlled, no general results are derived in this direction.
Later works, such as that of
\citet{SALP,BhatFaMo12:SALPNP},  repeat the analysis of \citet{CS}
in combinations with other ideas. However, no existing work that we know of addresses the above weaknesses of the result of \citet{CS}.

Our main contribution is a new suboptimality bound for the case when the constraint system is replaced with a smaller, 
linearly projected constraint system. We also propose a specific way of adding the extra constraint to keep the resulting LP bounded.
Rather than relying on combinatorial arguments (such as those at the heart of \citet{CS}), our argument uses previously unexploited geometric structure of the linear programs underlying MDPs.
As a result our bound avoids distribution-mismatch terms and we also remove the scaling with worst approximation error.
A specific outcome of our general result is the realization that it is beneficial to select states so that the ``feature vectors'' of all states when scaled with a fixed constant factor are included in the conic hull of the ``feature vectors'' underlying the selected states. This suggests to choose the basis functions so that this property can be satisfied by selecting only a few states. As we will argue, this property holds for several popular choices of basis functions.
A preliminary version of this paper without the theoretical analysis and without the geometric arguments was published in a short conference communication \cite{aaaipaper}.

\if0
